Computationally Efficient Simulation of Queues: The \proglang{R} Package \pkg{queuecomputer}

### 2018-11-06 main.tex
Manuscript style comments.

The queues we encounter in our everyday experience, where customers wait in line to be served by a server, is a useful analogy for many other processes. We say analogy because the word \textit{customers} could represent: MapReduce jobs \citep{lin_joint_2013}; patients in a hospital \citep{takagi_queueing_2016}; items in a manufacturing system \citep{dallery_manufacturing_1992}; calls to a call center \citep{gans_telephone_2003}; shipping containers in a seaport \citep{kozan_comparison_1997} or even cognitive tasks \citep{cao_queueing_2013}. Similarly \textit{server} could represent: a compute cluster; medical staff; machinery or a customer service representative at a call centre. Queueing systems can also be networked together to form \textit{queueing networks}. We can use queueing networks to build models of processes such as provision of internet services \citep{sutton_bayesian_2011}, passenger facilitation at international airports \citep{wu_review_2013} and emergency evacuations \citep{van_woensel_modeling_2007}. Clearly queueing systems and queueing networks are useful for understanding important real-world systems.

Performance measures for a given queueing system can often only be derived through simulation. Queues are usually simulated with discrete event simulation (DES) \citep[pg. 226]{insua2012bayesian}. In DES changes in state are discontinuous. The state is acted upon by a countable list of \textit{events} at certain times which cause the discontinuities. If the occurrence of an event is independent of everything except simulation time it is \textit{determined}, otherwise it is \textit{contingent} \citep{nance1981time}.

@@

A \textit{queueing system} is defined as follows. Each customer $i = 1,2,\cdots$ has an arrival time $a_i$ (or equivalently an inter-arrival time $\delta_i = a_i - a_{i-1}$, $a_0 = 0$) and an amount of time they require with a server, called the \textit{service time} $s_i$. Typically a server can serve only one customer at a time. A server which is currently serving another customer is said to be \textit{unavailable}, a server without a customer is \textit{available}. If all servers are unavailable when a customer arrives then customers must wait in the queue until a server is available. Detailed introductions to queueing systems can be found in standard texts such as \citet{bhat2015introduction}.

@@

Managers and planners are less interested in parameter inference and more interested in performance measures such as: $N(t)$, the number of customers in system at time $t$; $\bar{B}$, the average number of busy servers; $\rho$, the \textit{resource utilization}; and $\bar{\mathbf{w}}$, the average waiting time for customers. If $\lambda < K \mu$ the queueing system will eventually reach equilibrium and distributions of performance measures become independent of time.

@@

\citet{jackson_networks_1957} was one of the first to consider networks of queueing systems. In a \textit{Jackson network} there is a set of $J$ queueing systems. After a customer is served by queueing system $j$ they arrive at another queueing system with fixed probability $p_{j,k}$. Customers leave the system with probability $1 - \sum_{k=1}^{J} p_{j,k}$. Other examples of queueing networks include the \textit{tandem} \citep{glynn_departures_1991}, \textit{parallel} \citep{hunt_fast_1995} and the \textit{fork/join} \citep{kim_analysis_1989} topologies.

@@

Most models of queueing systems assume time-invariant inter-arrival and service processes. In practice many real-world queues have inter-arrival processes which are strongly time-dependent, such as: call-centres \citep{weinberg2007bayesian, brown2005statistical}, airport runways \citep{koopman1972air} and hospitals \citep{brahimi_queueing_1991}. In the case of the $M/M/1$ queue we can adapt the notation to $M(t)/M(t)/1$ to represent exponential  processes where parameters $\lambda(t)$ and $\mu(t)$ change with time. Such queueing systems are referred to as \textit{dynamic} queueing systems.

@@

This algorithm is simple and computationally efficient. At each iteration of the loop we need only search $\mathbf{b}$, a $K$ length vector for the minimum element in code line \ref{alg_line:search}. In the language of DES, we consider $\mathbf{b}$ as the system state and $\mathbf{a}$ as the event list, which are all \textit{determined} events. This differs from conventional DES approaches to modelling queueing systems where the queue length is the system state and both $\mathbf{a}$ and $\mathbf{d}$ constitute the event list, where the events of $\mathbf{a}$ are \textit{determined} and the events of $\mathbf{d}$ are continually updated and therefore \textit{contingent}.

@@

Suppose that the number of servers that customers can use changes throughout the day. This reflects realistic situations where more servers are rostered on for busier times of the day. We say that for a certain time $t$, the customers have a choice of $K(t)$ \textit{open} servers from $K$. This means that there are $K(t)$ servers rostered-on for time $t$. We define the term \textit{closed} as the opposite of open.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("queuecomputer")
R> arrivals <- cumsum(rexp(100))
R> head(arrivals)
\end{CodeInput}
\begin{CodeOutput}
[1] 0.693512 1.693399 2.425550 3.952405 3.961906 4.405492
\end{CodeOutput}
\begin{CodeInput}
R> service <- rexp(100)
R> departures <- queue_step(arrivals, service = service, servers = 2)
R> departures
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 100 × 6
   arrivals     service departures      waiting system_time server
      <dbl>       <dbl>      <dbl>        <dbl>       <dbl>  <dbl>
1  0.693512 0.830158956   1.523671 0.000000e+00 0.830158956      1
2  1.693399 0.817648174   2.511047 1.110223e-16 0.817648174      2
3  2.425550 0.002675641   2.428226 2.138047e-16 0.002675641      1
4  3.952405 0.667180991   4.619586 4.440892e-16 0.667180991      1
5  3.961906 0.551920432   4.513827 4.440892e-16 0.551920432      2
6  4.405492 1.069236762   5.583063 1.083341e-01 1.177570886      2
7  4.594253 1.110448926   5.730035 2.533279e-02 1.135781711      1
8  4.993053 0.766944956   6.350008 5.900099e-01 1.356954853      2
9  6.047412 0.805061421   6.852474 1.110223e-16 0.805061421      1
10 6.856338 1.317802131   8.174140 0.000000e+00 1.317802131      2
# ... with 90 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(departures)
\end{CodeInput}
\begin{CodeOutput}
Total customers:
 100
Missed customers:
 0
Mean waiting time:
 0.246
Mean response time:
 1.11
Utilization factor:
 0.53
Mean queue length:
 0.301
Mean number of customers in system:
 1.36
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> queuecomputer_output <- queue_step(arrivals = arrivals, 
+    service = service, servers = 2)
R> head(sort(depart(queuecomputer_output)))
\end{CodeInput}
\begin{CodeOutput}
[1] 1.340151 2.288112 2.639976 2.796572 3.249794 5.714967
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> simmer_output <- simmer_step(arrivals = arrivals, 
+    service = service, servers = 2)
R> head(simmer_output)
\end{CodeInput}
\begin{CodeOutput}
[1] 1.340151 2.288112 2.639976 2.796572 3.249794 5.714967
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
python> simpy_step(interarrivals, service)[0:6]
\end{CodeInput}
\begin{CodeOutput}
array([ 1.34015149,  2.28811237,  2.63997568,  2.79657232,  3.24979406, 
	5.7149671 ])
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> set.seed(1) 
R> n_customers <- 5e6
R> lambda <- 2
R> mu <- 1
R> interarrivals <- rexp(n_customers, lambda)
R> arrivals <- cumsum(interarrivals)
R> service <- rexp(n_customers, mu)
R> K = 3
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> MM3 <- queue_step(arrivals = arrivals, service = service, servers = K)
R> summary(MM3)
\end{CodeInput}
\begin{CodeOutput}
Total customers:
 5000000
Missed customers:
 0
Mean waiting time:
 0.445
Mean response time:
 1.44
Utilization factor:
 0.666
Mean queue length:
 0.889
Mean number of customers in system:
 2.89
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("queuecomputer")
R> library("randomNames") 
R> library("ggplot2")
R> set.seed(1) 
R> interarrivals <- rexp(20, 1)
R> arrivals <- cumsum(interarrivals)
R> customers <- randomNames(20, name.order = "first.last")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> service <- rexp(20, 0.5)
R> head(service)
\end{CodeInput}
\begin{CodeOutput}
[1] 2.4245155 1.1304372 0.3815756 0.5626325 2.0107932 1.4076141
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> queue_obj <- queue_step(arrivals, service, servers = 2, labels = customers)
R> head(queue_obj$departures_df)
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 6 × 7
   labels  arrivals   service departures      waiting system_time server
    <chr>     <dbl>     <dbl>      <dbl>        <dbl>       <dbl>  <dbl>
1 Kwabena 0.7551818 2.6669670   3.422149 0.000000e+00    2.666967      1
2 Beatriz 1.9368246 1.2434810   3.180306 0.000000e+00    1.243481      2
3 Dashawn 2.0825313 0.4197332   3.600039 1.097774e+00    1.517507      2
4  Karina 2.2223266 0.6188957   4.041045 1.199822e+00    1.818718      1
5    Ilea 2.6583952 2.2118725   5.811911 9.416435e-01    3.153516      2
6 Brianna 5.5533638 1.5483755   7.101739 2.220446e-16    1.548376      1
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> firstcustomers <- arrivals[1:2] + service[1:2]
R> firstcustomers
 \end{CodeInput}
 \begin{CodeOutput}
[1] 3.422149 3.180306
 \end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> firstcustomers[2] + service[3]
\end{CodeInput}
\begin{CodeOutput}
[1] 3.600039
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> depart(queue_obj)[1:3] - arrivals[1:3] - service[1:3]
\end{CodeInput}
\begin{CodeOutput}
[1] 0.000000 0.000000 1.097774
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> summary(queue_obj)
\end{CodeInput}
\begin{CodeOutput}
Total customers:
 20
Missed customers:
 0
Mean waiting time:
 1.15
Mean response time:
 3.69
Utilization factor:
 0.834
Mean queue length:
 0.858
Mean number of customers in system:
 2.42
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot(queue_obj, which = c(2, 4, 5, 6)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> Passenger_df
\end{CodeInput}
\begin{CodeOutput}
## # A tibble: 25,012 × 7
##                   ID FlightNo arrival  route_imm arrive_imm service_imm
##                <chr>   <fctr>   <dbl>     <fctr>      <dbl>       <dbl>
## 1     Cordell, Megan   ABI481  564.85     manual   566.8549  0.29075606
## 2    Matheson, Dylan   ABI481  564.85     manual   566.8532  0.15927226
## 3      Avitia, Renee   ABI481  564.85     manual   567.2014  0.22450319
## 4       Woods, Tyrel   ABI481  564.85 smart gate   566.8377  0.18222445
## 5   Pope, Christiana   ABI481  564.85 smart gate   566.0994  0.09031344
## 6   Espinoza, Mariah   ABI481  564.85 smart gate   566.8928  0.43900281
## 7  Pacheco, Charleen   ABI481  564.85     manual   567.5558  0.12917143
## 8    Harmon, Brendan   ABI481  564.85 smart gate   566.3114  0.30565961
## 9   William, Gerardo   ABI481  564.85 smart gate   567.2563  0.31975687
## 10       Hood, Colen   ABI481  564.85 smart gate   567.2181  0.33944458
## # ... with 25,002 more rows, and 1 more variables: bag_time <dbl>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> server_df <- data.frame(immigration_route = c("smart gate", "manual"))
R> server_df$servers <- 
+    list(5, as.server.stepfun(x = c(600,780), y = c(10,12,8)))
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> Passenger_df <- left_join(Passenger_df, server_df)
\end{CodeInput}
\begin{CodeOutput}
## Joining, by = "route_imm"
\end{CodeOutput}
\begin{CodeInput}
R> Passenger_df <- Passenger_df %>% 
+    group_by(route_imm) %>%
+    mutate(
+      departures_imm = 
+        queue(arrive_imm, service_imm, servers = servers[[1]])
+      ) %>% 
+    ungroup() %>% 
+    mutate(departures_bc = pmax.int(departures_imm, bag_time))
R> Passenger_df %>% 
+    select(FlightNo, arrive_imm, departures_imm, departures_bc)
\end{CodeInput}
\begin{CodeOutput}
## # A tibble: 25,012 × 4
##    FlightNo arrive_imm departures_imm departures_bc
##      <fctr>      <dbl>          <dbl>         <dbl>
## 1    ABI481   566.8549       578.8443      578.8443
## 2    ABI481   566.8532       578.6479      578.6479
## 3    ABI481   567.2014       579.6133      579.6133
## 4    ABI481   566.8377       571.9148      574.6535
## 5    ABI481   566.0994       570.2073      574.7161
## 6    ABI481   566.8928       572.2135      574.7912
## 7    ABI481   567.5558       579.9953      579.9953
## 8    ABI481   566.3114       570.6290      574.8944
## 9    ABI481   567.2563       573.3438      574.8996
## 10   ABI481   567.2181       573.1459      574.9295
## # ... with 25,002 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> Passenger_df %>% 
+    group_by(FlightNo, route_imm) %>%
+    summarise(
+      waiting_imm = mean(departures_imm - service_imm - arrive_imm), 
+      waiting_bc = mean(departures_bc - departures_imm)
+    )
\end{CodeInput}
\begin{CodeOutput}
## Source: local data frame [240 x 4]
## Groups: FlightNo [?]
## 
##    FlightNo  route_imm waiting_imm waiting_bc
##      <fctr>     <fctr>       <dbl>      <dbl>
## 1    ABI481     manual  11.2718493   6.292506
## 2    ABI481 smart gate   4.9586928  12.276112
## 3    AEB843     manual   0.8497289  16.594387
## 4    AEB843 smart gate   1.0133830  16.412237
## 5    ARH364     manual  12.5063642   3.795084
## 6    ARH364 smart gate   7.3569441   8.169201
## 7    BCH445     manual   1.8033024  13.700561
## 8    BCH445 smart gate   1.4368622  15.087956
## 9    BJN726     manual  19.5150170   2.064476
## 10   BJN726 smart gate   7.2110772   9.751714
## # ... with 230 more rows
\end{CodeOutput}
\begin{CodeInput}
R> Passenger_df %>% 
+    group_by(route_imm) %>%
+    summarise(
+      waiting_imm = mean(departures_imm - service_imm - arrive_imm), 
+      waiting_bc = mean(departures_bc - departures_imm)
+    )
\end{CodeInput}
\begin{CodeOutput}
## # A tibble: 2 × 3
##    route_imm waiting_imm waiting_bc
##       <fctr>       <dbl>      <dbl>
## 1     manual    8.252137   9.559913
## 2 smart gate    4.488549  12.786145
\end{CodeOutput}
\end{CodeChunk}

>>>>

The queues we encounter in our everyday experience, where customers wait in line to be served by a server, is a useful analogy for many other processes. We say analogy because the word customers could represent: MapReduce jobs \citep{lin_joint_2013}; patients in a hospital \citep{takagi_queueing_2016}; items in a manufacturing system \citep{dallery_manufacturing_1992}; calls to a call center \citep{gans_telephone_2003}; shipping containers in a seaport \citep{kozan_comparison_1997} or even cognitive tasks \citep{cao_queueing_2013}. Similarly server could represent: a compute cluster; medical staff; machinery or a customer service representative at a call centre. Queueing systems can also be networked together to form queueing networks. We can use queueing networks to build models of processes such as provision of internet services \citep{sutton_bayesian_2011}, passenger facilitation at international airports \citep{wu_review_2013} and emergency evacuations \citep{van_woensel_modeling_2007}. Clearly queueing systems and queueing networks are useful for understanding important real-world systems.

Performance measures for a given queueing system can often only be derived through simulation. Queues are usually simulated with discrete event simulation (DES) \citep[pg. 226]{insua2012bayesian}. In DES changes in state are discontinuous. The state is acted upon by a countable list of events at certain times which cause the discontinuities. If the occurrence of an event is independent of everything except simulation time it is determined, otherwise it is contingent \citep{nance1981time}.

@@

A queueing system is defined as follows. Each customer $i = 1,2,\cdots$ has an arrival time $a_i$ (or equivalently an inter-arrival time $\delta_i = a_i - a_{i-1}$, $a_0 = 0$) and an amount of time they require with a server, called the service time $s_i$. Typically a server can serve only one customer at a time. A server which is currently serving another customer is said to be unavailable, a server without a customer is available. If all servers are unavailable when a customer arrives then customers must wait in the queue until a server is available. Detailed introductions to queueing systems can be found in standard texts such as \citet{bhat2015introduction}.

@@

Managers and planners are less interested in parameter inference and more interested in performance measures such as: $N(t)$, the number of customers in system at time $t$; $\bar{B}$, the average number of busy servers; $\rho$, the resource utilization; and $\bar{\mathbf{w}}$, the average waiting time for customers. If $\lambda < K \mu$ the queueing system will eventually reach equilibrium and distributions of performance measures become independent of time.

@@

\citet{jackson_networks_1957} was one of the first to consider networks of queueing systems. In a Jackson network there is a set of $J$ queueing systems. After a customer is served by queueing system $j$ they arrive at another queueing system with fixed probability $p_{j,k}$. Customers leave the system with probability $1 - \sum_{k=1}^{J} p_{j,k}$. Other examples of queueing networks include the tandem \citep{glynn_departures_1991}, parallel \citep{hunt_fast_1995} and the fork/join \citep{kim_analysis_1989} topologies.

@@

Most models of queueing systems assume time-invariant inter-arrival and service processes. In practice many real-world queues have inter-arrival processes which are strongly time-dependent, such as: call-centres \citep{weinberg2007bayesian, brown2005statistical}, airport runways \citep{koopman1972air} and hospitals \citep{brahimi_queueing_1991}. In the case of the $M/M/1$ queue we can adapt the notation to $M(t)/M(t)/1$ to represent exponential  processes where parameters $\lambda(t)$ and $\mu(t)$ change with time. Such queueing systems are referred to as dynamic queueing systems.

@@

This algorithm is simple and computationally efficient. At each iteration of the loop we need only search $\mathbf{b}$, a $K$ length vector for the minimum element in code line \ref{alg_line:search}. In the language of DES, we consider $\mathbf{b}$ as the system state and $\mathbf{a}$ as the event list, which are all determined events. This differs from conventional DES approaches to modelling queueing systems where the queue length is the system state and both $\mathbf{a}$ and $\mathbf{d}$ constitute the event list, where the events of $\mathbf{a}$ are determined and the events of $\mathbf{d}$ are continually updated and therefore contingent.

@@

Suppose that the number of servers that customers can use changes throughout the day. This reflects realistic situations where more servers are rostered on for busier times of the day. We say that for a certain time $t$, the customers have a choice of $K(t)$ open servers from $K$. This means that there are $K(t)$ servers rostered-on for time $t$. We define the term closed as the opposite of open.

@@

\begin{CodeChunk}
\begin{Sinput}
R> library("queuecomputer")
R> arrivals <- cumsum(rexp(100))
R> head(arrivals)
\end{Sinput}
\begin{Soutput}
[1] 0.693512 1.693399 2.425550 3.952405 3.961906 4.405492
\end{Soutput}
\begin{Sinput}
R> service <- rexp(100)
R> departures <- queue_step(arrivals, service = service, servers = 2)
R> departures
\end{Sinput}
\begin{Soutput}
# A tibble: 100 × 6
   arrivals     service departures      waiting system_time server
      <dbl>       <dbl>      <dbl>        <dbl>       <dbl>  <dbl>
1  0.693512 0.830158956   1.523671 0.000000e+00 0.830158956      1
2  1.693399 0.817648174   2.511047 1.110223e-16 0.817648174      2
3  2.425550 0.002675641   2.428226 2.138047e-16 0.002675641      1
4  3.952405 0.667180991   4.619586 4.440892e-16 0.667180991      1
5  3.961906 0.551920432   4.513827 4.440892e-16 0.551920432      2
6  4.405492 1.069236762   5.583063 1.083341e-01 1.177570886      2
7  4.594253 1.110448926   5.730035 2.533279e-02 1.135781711      1
8  4.993053 0.766944956   6.350008 5.900099e-01 1.356954853      2
9  6.047412 0.805061421   6.852474 1.110223e-16 0.805061421      1
10 6.856338 1.317802131   8.174140 0.000000e+00 1.317802131      2
# ... with 90 more rows
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> summary(departures)
\end{Sinput}
\begin{Soutput}
Total customers:
 100
Missed customers:
 0
Mean waiting time:
 0.246
Mean response time:
 1.11
Utilization factor:
 0.53
Mean queue length:
 0.301
Mean number of customers in system:
 1.36
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> queuecomputer_output <- queue_step(arrivals = arrivals, 
+    service = service, servers = 2)
R> head(sort(depart(queuecomputer_output)))
\end{Sinput}
\begin{Soutput}
[1] 1.340151 2.288112 2.639976 2.796572 3.249794 5.714967
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> simmer_output <- simmer_step(arrivals = arrivals, 
+    service = service, servers = 2)
R> head(simmer_output)
\end{Sinput}
\begin{Soutput}
[1] 1.340151 2.288112 2.639976 2.796572 3.249794 5.714967
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
python> simpy_step(interarrivals, service)[0:6]
\end{Sinput}
\begin{Soutput}
array([ 1.34015149,  2.28811237,  2.63997568,  2.79657232,  3.24979406, 
	5.7149671 ])
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> set.seed(1) 
R> n_customers <- 5e6
R> lambda <- 2
R> mu <- 1
R> interarrivals <- rexp(n_customers, lambda)
R> arrivals <- cumsum(interarrivals)
R> service <- rexp(n_customers, mu)
R> K = 3
\end{Sinput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> MM3 <- queue_step(arrivals = arrivals, service = service, servers = K)
R> summary(MM3)
\end{Sinput}
\begin{Soutput}
Total customers:
 5000000
Missed customers:
 0
Mean waiting time:
 0.445
Mean response time:
 1.44
Utilization factor:
 0.666
Mean queue length:
 0.889
Mean number of customers in system:
 2.89
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> library("queuecomputer")
R> library("randomNames") 
R> library("ggplot2")
R> set.seed(1) 
R> interarrivals <- rexp(20, 1)
R> arrivals <- cumsum(interarrivals)
R> customers <- randomNames(20, name.order = "first.last")
\end{Sinput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> service <- rexp(20, 0.5)
R> head(service)
\end{Sinput}
\begin{Soutput}
[1] 2.4245155 1.1304372 0.3815756 0.5626325 2.0107932 1.4076141
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> queue_obj <- queue_step(arrivals, service, servers = 2, 
+    labels = customers)
R> head(queue_obj$departures_df)
\end{Sinput}
\begin{Soutput}
# A tibble: 6 × 7
   labels  arrivals   service departures      waiting system_time server
    <chr>     <dbl>     <dbl>      <dbl>        <dbl>       <dbl>  <dbl>
1 Kwabena 0.7551818 2.6669670   3.422149 0.000000e+00    2.666967      1
2 Beatriz 1.9368246 1.2434810   3.180306 0.000000e+00    1.243481      2
3 Dashawn 2.0825313 0.4197332   3.600039 1.097774e+00    1.517507      2
4  Karina 2.2223266 0.6188957   4.041045 1.199822e+00    1.818718      1
5    Ilea 2.6583952 2.2118725   5.811911 9.416435e-01    3.153516      2
6 Brianna 5.5533638 1.5483755   7.101739 2.220446e-16    1.548376      1
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> firstcustomers <- arrivals[1:2] + service[1:2]
R> firstcustomers
 \end{Sinput}
 \begin{Soutput}
[1] 3.422149 3.180306
 \end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> firstcustomers[2] + service[3]
\end{Sinput}
\begin{Soutput}
[1] 3.600039
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> depart(queue_obj)[1:3] - arrivals[1:3] - service[1:3]
\end{Sinput}
\begin{Soutput}
[1] 0.000000 0.000000 1.097774
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> summary(queue_obj)
\end{Sinput}
\begin{Soutput}
Total customers:
 20
Missed customers:
 0
Mean waiting time:
 1.15
Mean response time:
 3.69
Utilization factor:
 0.834
Mean queue length:
 0.858
Mean number of customers in system:
 2.42
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> plot(queue_obj, which = c(2, 4, 5, 6)
\end{Sinput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> Passenger_df
\end{Sinput}
\begin{Soutput}
# A tibble: 25,012 × 7
                  ID FlightNo arrival  route_imm arrive_imm service_imm
               <chr>   <fctr>   <dbl>     <fctr>      <dbl>       <dbl>
1     Cordell, Megan   ABI481  564.85     manual   566.8549  0.29075606
2    Matheson, Dylan   ABI481  564.85     manual   566.8532  0.15927226
3      Avitia, Renee   ABI481  564.85     manual   567.2014  0.22450319
4       Woods, Tyrel   ABI481  564.85 smart gate   566.8377  0.18222445
5   Pope, Christiana   ABI481  564.85 smart gate   566.0994  0.09031344
6   Espinoza, Mariah   ABI481  564.85 smart gate   566.8928  0.43900281
7  Pacheco, Charleen   ABI481  564.85     manual   567.5558  0.12917143
8    Harmon, Brendan   ABI481  564.85 smart gate   566.3114  0.30565961
9   William, Gerardo   ABI481  564.85 smart gate   567.2563  0.31975687
10       Hood, Colen   ABI481  564.85 smart gate   567.2181  0.33944458
# ... with 25,002 more rows, and 1 more variables: bag_time <dbl>
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> server_df <- data.frame(immigration_route = c("smart gate", "manual"))
R> server_df$servers <- 
+    list(5, as.server.stepfun(x = c(600,780), y = c(10,12,8)))
\end{Sinput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> Passenger_df <- left_join(Passenger_df, server_df, by = "route_imm")
R> Passenger_df <- Passenger_df %>% 
+    group_by(route_imm) %>%
+    mutate(
+      departures_imm = 
+        queue(arrive_imm, service_imm, servers = servers[[1]])
+      ) %>% 
+    ungroup() %>% 
+    mutate(departures_bc = pmax.int(departures_imm, bag_time))
R> Passenger_df %>% 
+    select(FlightNo, arrive_imm, departures_imm, departures_bc)
\end{Sinput}
\begin{Soutput}
# A tibble: 25,012 × 4
   FlightNo arrive_imm departures_imm departures_bc
     <fctr>      <dbl>          <dbl>         <dbl>
1    ABI481   566.8549       578.8443      578.8443
2    ABI481   566.8532       578.6479      578.6479
3    ABI481   567.2014       579.6133      579.6133
4    ABI481   566.8377       571.9148      574.6535
5    ABI481   566.0994       570.2073      574.7161
6    ABI481   566.8928       572.2135      574.7912
7    ABI481   567.5558       579.9953      579.9953
8    ABI481   566.3114       570.6290      574.8944
9    ABI481   567.2563       573.3438      574.8996
10   ABI481   567.2181       573.1459      574.9295
# ... with 25,002 more rows
\end{Soutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{Sinput}
R> Passenger_df %>% 
+    group_by(FlightNo, route_imm) %>%
+    summarise(
+      waiting_imm = mean(departures_imm - service_imm - arrive_imm), 
+      waiting_bc = mean(departures_bc - departures_imm)
+    )
\end{Sinput}
\begin{Soutput}
Source: local data frame [240 x 4]
Groups: FlightNo [?]

   FlightNo  route_imm waiting_imm waiting_bc
     <fctr>     <fctr>       <dbl>      <dbl>
1    ABI481     manual  11.2718493   6.292506
2    ABI481 smart gate   4.9586928  12.276112
3    AEB843     manual   0.8497289  16.594387
4    AEB843 smart gate   1.0133830  16.412237
5    ARH364     manual  12.5063642   3.795084
6    ARH364 smart gate   7.3569441   8.169201
7    BCH445     manual   1.8033024  13.700561
8    BCH445 smart gate   1.4368622  15.087956
9    BJN726     manual  19.5150170   2.064476
10   BJN726 smart gate   7.2110772   9.751714
# ... with 230 more rows
\end{Soutput}
\begin{Sinput}
R> Passenger_df %>% 
+    group_by(route_imm) %>%
+    summarise(
+      waiting_imm = mean(departures_imm - service_imm - arrive_imm), 
+      waiting_bc = mean(departures_bc - departures_imm)
+    )
\end{Sinput}
\begin{Soutput}
# A tibble: 2 × 3
   route_imm waiting_imm waiting_bc
      <fctr>       <dbl>      <dbl>
1     manual    8.252137   9.559913
2 smart gate    4.488549  12.786145
\end{Soutput}
\end{CodeChunk}
